Large-scale feature selection with Gaussian mixture models for the classification of high dimensional remote sensing images

### 2017-01-11 response.tex

\begin{revbox}
  Following the question 3, why choosing the optimal number of variables based on the increase of criterion instead of the highest value?
  \begin{resbox}
    We also considered this possibility and it is a choice available in the code of the OTB module we provide. However, selecting the maximum value of the criteria without considering the slope of the curve might be hazardous. At some point, adding more feature does not improve significantly the discrimination of the classes, until a too large number of variables is reached and statistical issues arise. It is therefore safer to stop when the increase of criteria stops to be significant.
\end{resbox}
\end{revbox}

% \begin{figure}[!t]
%   \centering
%   \begin{tikzpicture}
%     \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
%       \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
%     \end{axis};
%   \end{tikzpicture}
%   \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Potsdam dataset with 1000 samples by class.\label{fig:crit-evol}}
% \end{figure}

\begin{revbox}
  In the experiment with Aisa dataset, although several numbers of training samples are tested, the number of training samples seems carefully selected by the authors to show the cases where the GMM classifiers perform better. I guess when a larger number of samples are trained, the random forest will have a higher classification rate. And when a smaller number of samples are used, the KNN will have a shorter training time. The GMM only wins in the classification time regardless of the sample number. It’ll be good to show these cases.
  \begin{resbox}
    The reviewer is right regarding the number of training samples. GMM (not restricted to the feature selection case) does not benefit as much as other machine learning algorithms of the increase of the number of training samples. For the feature selection case, it is discussed in ref [27] for instance, with KNN, RF and SVM. Please note it is also discussed in the paper for the Potsdam dataset.

@@

    \begin{table}[!t]
        \centering
        \caption{Average classification accuracy for 1,000 samples by class averaged over 5 trials  evaluated with overall accuracy and mean of F1-score (standard deviation in parenthesis).\label{tab:potsdam-otbsimu-othereval}}
        \begin{tabular}{lcccccc}\toprule
             & \multicolumn{2}{c}{\bfseries 5\_11 (train)} & \multicolumn{2}{c}{\bfseries 5\_12 (test)} & \multicolumn{2}{c}{\bfseries 3\_10 (test)} \\ \cmidrule{2-7}
             & {\bfseries OA} & {\bfseries F1-mean} & {\bfseries OA} & {\bfseries F1-mean} & {\bfseries OA} & {\bfseries F1-mean} \\ \cmidrule{2-7}

            GMM SFS JM &      &&&&&\\
            GMM SFS divKL &   &&&&&\\
            GMM SFS kappa &   &&&&&\\
            GMM SFS OA &      &&&&&\\
            GMM SFS F1mean &  &&&&&\\
            GMM SFFS JM &     &&&&&\\
            GMM SFFS divKL &  &&&&&\\
            GMM SFFS kappa &  &&&&&\\
            GMM SFFS OA &     &&&&&\\
            GMM SFFS F1mean & &&&&&\\
            \bottomrule
        \end{tabular}
    \end{table}

\begin{revbox}
  According to Table IV in page 16. GMM SFS JM outperforms GMM SFS kappa a little bit under the kappa criterion. Since GMM SFS kappa updates the features according to the kappa criterion while GMM SFS JM updates the features according to the JM criterion, shouldn’t GMM SFS kappa outperforms GMM SFS JM under the kappa criterion? Or is there some explanation to it?
  \begin{resbox}
    We would like to thank the reviewer for this important comment. First, please note that the score difference between GMM SFS JM and GMM SFS kappa in Table IV is not statistically significant according to the Wilcoxon rank-sum test. Second, the Kappa is estimated by cross-validation during the feature selection, which is a  biased estimators and therefore, it is possible to observe significant differences between the estimated kappa and the true Kappa (usually, a lower value), see ref [36] Chapter 7.

    Our guess is that if the Gaussian hypothesis is a good approximation for the used data, the JM metric performs well  with  performances similar to the ones obtained with the Kappa. It is the case for the AISA data set, while for the Potsdam data, the Kappa coefficient outperformed the JM metric. Our recommendation would be to start with JM since it is fast, and then try Kappa if necessary.
  \end{resbox}
\end{revbox}

@@

\begin{revbox}
  In Table VI and Table VII, the \# of selected features are different for different proposed methods. And the classification performance drops when the \# of selected features decreases. According to Algorithm 2 and Algorithm 3, the \# of selected features should be equal to maxVarNb. So why the \# of selected features are different for different proposed methods? The comparison is not fair to me if the \# of selected features are different. The authors should also show the \# of selected features in Table IV or Table V as well.
  \begin{resbox}
    We apologize  for the misunderstanding,  this part was  unclear in
    the primary  version of  the paper. The  \emph{maxVarNb} parameter
    corresponds  to  the number  of  bands  searched by  the  iterative
    algorithm. To select  the optimal number of  selected features, the
    criterion  evolution is  analyzed, as  explain in  the end  of the
    Section VI.A. The optimal number  of retained features is found by
    computing  the discrete  derivative  of the  criterion between  two
    iterations and normalizing it by  its maximum value. The number of
    selected  features corresponds  to the  last iteration  before the
    value drops below $10^{-3}$ for  all datasets. See response to the
    Reviewer 1 (1.3 and 1.4).

    Therefore, in practice, the optimal number can be different from one method to another.  We believe that it is fair to use a different number of variables for each method because it also reflects the ability of a criterion to identify relevant features.

    The Table V has been modified to include the \# of selected features.
  \end{resbox}
\end{revbox}

\begin{revbox}
  A reference or an explanation should be made in Page 1, line 36 regarding the curse of dimensionality
  \begin{resbox}
    We added the following reference: \emph{D. Donoho, ``High-dimensional data analysis: the curses and blessing of dimensionality,'' in AMS Mathematical challenges of the 21st century, 2000}.
  \end{resbox}
\end{revbox}

@@

\section{Reviewer 3}

\begin{revbox}
  Although the method seems promising, there are some significant issues the authors should address to. The key novelty and contribution of this technique is unfortunately unclear. The authors dedicate a lot of space in explaining already known mathematical formulations of the Gaussian Mixture Models and the feature selection techniques, in comparison with their proposed scheme.
  \begin{resbox}
    We totally agree that some mathematical formulations are already known. Nevertheless, we think that these formulations in this particular context has not been exploited so far. The results of our experiment shows that the proposed algorithm has the advantage to be very efficient in term of processing time with a fairly reasonable cost in term of classification accuracy. It has to be noticed that compared to similar methods using Gaussian Mixture model, no additional hypothesis are made on the covariance matrix which is tractable only because of the efficient implementation.
  \end{resbox}
\end{revbox}

### 2017-01-12 response.tex

    The Table V has been modified to include the \# of selected features.
  \end{resbox}
\end{revbox}

\begin{revbox}
  A reference or an explanation should be made in Page 1, line 36 regarding the curse of dimensionality
  \begin{resbox}
    We added the following reference: \emph{D. Donoho, ``High-dimensional data analysis: the curses and blessing of dimensionality,'' in AMS Mathematical challenges of the 21st century, 2000}.
  \end{resbox}
\end{revbox}

\section{Reviewer 3}

\begin{revbox}
  Although the method seems promising, there are some significant issues the authors should address to. The key novelty and contribution of this technique is unfortunately unclear. The authors dedicate a lot of space in explaining already known mathematical formulations of the Gaussian Mixture Models and the feature selection techniques, in comparison with their proposed scheme.
  \begin{resbox}
    We totally agree that some mathematical formulations are conventional. We believe that it is necessary to provide to readers sufficient information to explain clearly the contribution of the proposed paper. The contributions are based on specific derivation of the GMM and the FE formulation. Without these primary recall, we are afraid that readers can be lost. Concerning the contribution of the paper, they are stated in the antepenultimate paragraph of the introduction. They concern the expression of efficient updates rules using block matrix processing. Furthermore, an open source C++ library is made available for the community. The results of our experiment shows that the proposed algorithm has the advantage to be very efficient in term of processing time with a fairly reasonable cost in term of classification accuracy. It has to be noticed that compared to similar methods using Gaussian Mixture model, no additional hypothesis are made on the covariance matrix which is tractable only because of the efficient implementation.

    To make the contribution more clear, the antepenultimate paragraph has been modified to

\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~\cite{fauvel2015fast}. An smart implementation of the feature selection updates rules are presented in order to perform efficiently on large amount of data. They are based on linear algebra on block matrix applied to the covariance matrix of the conditional class density function.  Furthermore, a floating version of the algorithm is proposed and evaluated. Several \emph{goodness of fit} criteria are proposed to handle unbalanced training set, extending the conventional \emph{overall accuracy} measure.  Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~[13].}
  \end{resbox}
\end{revbox}

@@

\begin{revbox}
  Additionally, critical references are missing, such as:
  \begin{enumerate}
  \item K. Karalas, G. Tsagkatakis, M. Zervakis, and P. Tsakalides, "Land Classification Using Remotely Sensed Data: Going Multi-Label," IEEE Transactions on Geoscience and Remote Sensing, 2016, doi: 10.1109/TGRS.2016.2520203
  \item K. Karalas, G. Tsagkatakis, M. Zervakis, and P. Tsakalides. "Deep learning for multi-label land cover classification", in SPIE Remote Sensing, Toulouse
  \end{enumerate}

  \begin{resbox}
    We thank the reviewer for these two interesting references. However, after reading them with interest, we think the issues address in these papers are somewhat unrelated to the feature selection proposed in this work.  First reference reviews ways to assign several labels to one sample and so performs spectral unmixing. Second reference focus on the same topic but develop a deep learning approach. Although very interesting, we do not address the subject of multi-labeling in our work.  According to these observations and to keep the paper short, we choose to not include these references in the revised paper.
  \end{resbox}
\end{revbox}

\begin{revbox}
  Another drawback is that the English language is poorly written, since it contains several critical syntax and typographical errors. As a result, some parts of the text are unclarified and misunderstood.
  \begin{resbox}
    We proofread the article to correct as many errors as possible. If the reviewer still see weird sentences or typo, we will happy pleased to modify or correct them.

### 2017-01-12 paper.tex
update
Réponse au reviewer III
- Modif des réponses au Rev III
- Modif de l'avant derniere paragaph sur les contributions

In the spectral or temporal domain, a pixel is represented by a vector
for   which  each   component  corresponds   to  a   spectral/temporal
measurement.   The size  of  the  vector is  therefore  the number  of
spectral  or temporal  measurements.  For  hyperspectral images,  this
number is  typically about several  hundreds while for  the Sentinel-2
multitemporal images, the number of spectro-temporal measurement for a
given  year  is approximately  one  thousand.   When working  in  high
dimensional  space,  statistical  methods  made for  low  or  moderate
dimensional  space do  not  adapt  well.  For  instance,  the rate  of
convergence of the statistical estimation decreases when the dimension
grows while conjointly the number of parameters to estimate increases,
making    the    estimation    of   the    model    parameters    very
difficult~\cite{donoho}.  Consequently,  with a limited  training set,
beyond a certain limit, the classification accuracy actually decreases
as the number of features increases~\cite{hughes}.  For the purpose of
classification,  these  problems are  related  to  the \emph{curse  of
  dimensionality}~\cite{hughes}. This  is a  major drawback  in many  remote sensing
application since it  is difficult to collect a large  and an accurate
ground-truth.   An intensive  work has  been performed  in the  remote
sensing community  to build accurate classifiers  for high dimensional
images.  Bayesian models~\cite{book:landgrebe}, feature extraction and
feature   reduction   techniques~\cite{book:landgrebe,DR:guided:tour},
random   forest~\cite{1396322},  neural   networks~\cite{5411821}  and
kernel methods~\cite{kernel:methods:rs} have been investigated for the
classification of such images.

@@

The contribution of the paper is a forward feature selection method using GMM in continuation of~\cite{fauvel2015fast}. An efficient implementation of the method, using linear algebra on block matrix, is presented in order to handle large amount of data. Furthermore, a floating version of the algorithm is proposed. Several criteria are proposed to handle unbalanced training set.  Finally, the developed algorithm is made available as a remote module of the Orfeo Toolbox~\cite{christophe2008orfeo}.

>>>>

In the spectral or temporal domain, a pixel is represented by a vector
for   which  each   component  corresponds   to  a   spectral/temporal
measurement.   The size  of  the  vector is  therefore  the number  of
spectral  or temporal  measurements.  For  hyperspectral images,  this
number is  typically about several  hundreds while for  the Sentinel-2
multitemporal images, the number of spectro-temporal measurement for a
given  year  is approximately  one  thousand.   When working  in  high
dimensional  space,  statistical  methods  made for  low  or  moderate
dimensional  space do  not  adapt  well.  For  instance,  the rate  of
convergence of the statistical estimation decreases when the dimension
grows while conjointly the number of parameters to estimate increases,
making    the    estimation    of   the    model    parameters    very
difficult~\cite{donoho}.  Consequently,  with a limited  training set,
beyond a certain limit, the classification accuracy actually decreases
as the number of features increases~\cite{hughes}.  For the purpose of
classification,  these  problems are  related  to  the \emph{curse  of
  dimensionality}~\cite{donoho}.  This is  a  major  drawback in  many
remote sensing  application since it  is difficult to collect  a large
and an accurate ground-truth.  An intensive work has been performed in
the remote  sensing community to  build accurate classifiers  for high
dimensional  images.   Bayesian models~\cite{book:landgrebe},  feature
extraction              and             feature              reduction
techniques~\cite{book:landgrebe,DR:guided:tour},                random
forest~\cite{1396322},   neural  networks~\cite{5411821}   and  kernel
methods~\cite{kernel:methods:rs}  have   been  investigated   for  the
classification of such images.

@@

\rev{The contribution of the paper is a extension of the forward feature selection method proposed in~\cite{fauvel2015fast}. An smart implementation of the feature selection updates rules are presented in order to perform efficiently on large amount of data. They are based on linear algebra on block matrix applied to the covariance matrix of the conditional class density function.  Furthermore, a floating version of the algorithm is proposed and evaluated. Several \emph{goodness of fit} criteria are proposed to handle unbalanced training set, extending the conventional \emph{overall accuracy} measure.  Finally, the developed algorithm is made available to the scientific community as a remote module of the Orfeo Toolbox~\cite{christophe2008orfeo}.}
